Cross-Validation Default

### 2023-06-28 minos/validation/cross_validation_default.Rmd
Added some new functions to the utils_validation_vis.R, and created a cross_validation_default.Rmd to plot results from 5-fold cross-validation

---
title: "Cross Validation Analysis"
output: html_notebook
---

This notebook will compare the results of the cross-validation simulation run with real data from Understanding Society, so that we can evaluate the transition models.

# Method

After generation of the final datasets (final_US), the respondents were separated into 2 roughly equal halves based on their pidp. One half was labelled as the transition dataset, and the other for simulation. Transition models were then fit using the transition dataset, with the other half then simulated with these models.

The results of the simulation will now be compared to the simulation dataset to see how closely our models mimick reality.

@@

```{r, include=FALSE}
# cross validation simulation output
#out.path <- 'Minos/output/cross_validation/'
out.path <- here::here('output', 'cross_validation_default/')
cv.dat <- read_singular_local_out(out.path, 'baseline')

# cv simulation pop (not used for transitions)
cv.simul.files <- list.files(here::here('data', 'final_US', 'cross_validation', 'simulation'),
                             pattern = '[0-9]{4}_US_cohort.csv',
                             full.names = TRUE)
raw.dat <- do.call(rbind, lapply(cv.simul.files, read.csv))
```

# Analysis

@@

```{r}
# pidp filter
cv.pidps <- cv.dat %>% select(pidp) %>% unique()
raw.pidps <- raw.dat %>% select(pidp) %>% unique()
both.pidps <- intersect(cv.pidps, raw.pidps)

# year filter
cv.years <- cv.dat %>% select(time) %>% unique()
cv.years <- cv.years[['time']]

cv <- cv.dat %>% filter(pidp %in% both.pidps$pidp)
raw <- raw.dat %>% filter(pidp %in% both.pidps$pidp, 
                      time %in% cv.years)

rm(cv.pidps, raw.pidps, both.pidps, cv.years)
```

## hh_income

@@

plot_mean_comparison(pivoted.df = income.pivoted, 
                     var = 'hh_income', 
                     group.var = 'scenario',
                     save = TRUE,
                     save.path = '/home/luke/Documents/WORK/MINOS/VALIDATION_PLOTS/Cross-Validation/')

@@

### Snapshots

Going to produce some comparison plots for snapshots throughout the simulation.

```{r}
snap.timelist <- c(2014, 2016, 2018, 2020)

cv.snap <- cv %>%
  select(pidp, time, hh_income) %>%
  filter(time %in% snap.timelist) %>%
  rename(predicted = hh_income)

raw.snap <- raw %>%
  select(pidp, time, hh_income) %>%
  filter(time %in% snap.timelist) %>%
  rename(observed = hh_income)

snap <- merge(cv.snap, raw.snap,
                   by = c('pidp', 'time'))

asinh_trans <- scales::trans_new(
  "inverse_hyperbolic_sine",
  transform = function(x) {asinh(x)},
  inverse = function(x) {sinh(x)}
)

ggplot(snap, aes(x = observed, y = predicted)) +
  geom_point() +
  scale_y_continuous(trans=asinh_trans) +
  scale_x_continuous(trans=asinh_trans) +
  geom_abline(slope=1, intercept=0) +
  stat_ellipse(color='red') +
  facet_wrap(~time) +
  labs(title = 'hh_income observed vs predicted',
       subtitle = 'inverse hyperbolic sine transformation')

rm(snap.timelist, cv.snap, raw.snap, snap)
```

@@

### T-tests

```{r}
income.t.tests <- two_sample_t_test(cv, 'cv', raw, 'raw', 'hh_income')
income.t.tests
```

```{r}
yearly_box_plots(pivoted.df = income.pivoted,
                 var = 'hh_income',
                 scen1 = 'raw',
                 scen2 = 'simulation',
                 subset.max = 5000,
                 subset.min = -5000)
```

## SF12

```{r}
sf12.pivoted <- combine_and_pivot_long(df1 = cv, 
                                       df1.name = 'simulated', 
                                       df2 = raw, 
                                       df2.name = 'raw', 
                                       var = 'SF_12')

plot_mean_comparison(pivoted.df = sf12.pivoted, 
                     var = 'SF_12', 
                     group.var = 'scenario',
                     save = TRUE,
                     save.path = '/home/luke/Documents/WORK/MINOS/VALIDATION_PLOTS/Cross-Validation/')

compare_boxplot_long(pivoted.df = sf12.pivoted,
                     var = 'SF_12',
                     scen1 = 'raw',
                     scen2 = 'simulation')
```

### Snapshots

```{r}
observed <- raw
predicted <- cv
var <- 'SF_12'
target.year <- 2020

# get just one year
o.end <- observed %>% filter(time == target.year, .data[[var]] != -8.0)
p.end <- predicted %>% filter(time == target.year, .data[[var]] != -8.0)
# get just the SF12 columns and rename for later
o.end <- o.end %>% select(pidp, .data[[var]]) %>% rename(observed = var)
p.end <- p.end %>% select(pidp, .data[[var]]) %>% rename(predicted = var)

# combine before we plot
combined <- merge(o.end, p.end, by = 'pidp')

p <- ggplot(data = combined, aes(x = observed, y = predicted)) +
  geom_point(alpha = 0.6, size=0.1) +
  geom_smooth() +
  geom_abline(intercept = 0, scale=1) +
  theme(legend.position = c(0.15, 0.9)) +
  labs(title = paste0(var, ' - ', target.year))
p

p1 <- ggMarginal(p, type='densigram', xparams = list(position = 'dodge'), yparams=list(position = 'dodge'))
p1
```

```{r}
snap.timelist <- c(2014, 2016, 2018, 2020)

cv.snap <- cv %>%
  select(pidp, time, SF_12) %>%
  filter(time %in% snap.timelist) %>%
  rename(predicted = SF_12)

raw.snap <- raw %>%
  select(pidp, time, SF_12) %>%
  filter(time %in% snap.timelist) %>%
  rename(observed = SF_12)

snap <- merge(cv.snap, raw.snap,
                   by = c('pidp', 'time'))
snap$time <- as.factor(snap$time)

ggplot(snap, aes(x = observed, y = predicted, group = time)) +
  geom_point(size = 1) +
  geom_abline(slope=1, intercept=0) +
  stat_ellipse(color='red') +
  facet_wrap(~time) +
  labs(title = 'SF_12')

rm(snap.timelist, cv.snap, raw.snap, snap)
```

@@

### T-Tests

```{r}
sf12.t.tests <- two_sample_t_test(cv, 'cv', raw, 'raw', 'hh_income')
sf12.t.tests
```

```{r}
yearly_box_plots(pivoted.df = sf12.pivoted,
                 var = 'SF_12',
                 scen1 = 'raw',
                 scen2 = 'simulation')
```

@@

```{r}
yearly_box_plots(pivoted.df = nut.pivoted,
                 var = 'nutrition_quality',
                 scen1 = 'raw',
                 scen2 = 'simulation')
```

@@

```{r}
yearly_box_plots(pivoted.df = tobacco.pivoted,
                 var = 'ncigs',
                 scen1 = 'raw',
                 scen2 = 'simulation',
                 subset.max = 100,
                 subset.min = 0)
```

cv_ordinal_plots(pivoted.df = hous.pivoted, 
                 var = 'housing_quality',
                 save = TRUE,
                 save.path = '/home/luke/Documents/WORK/MINOS/VALIDATION_PLOTS/Cross-Validation/')
```

cv_ordinal_plots(pivoted.df = nhsafe.pivoted, 
                 var = 'neighbourhood_safety',
                 save = TRUE,
                 save.path = '/home/luke/Documents/WORK/MINOS/VALIDATION_PLOTS/Cross-Validation/')
```

cv_ordinal_plots(pivoted.df = lnly.pivoted, 
                 var = 'loneliness',
                 save = TRUE,
                 save.path = '/home/luke/Documents/WORK/MINOS/VALIDATION_PLOTS/Cross-Validation/')
```

>>>>

---
title: "Cross-Validation Default"
output: html_notebook
---

@@

```{r, include=FALSE}
# cross validation simulation output
#out.path <- 'Minos/output/cross_validation/'
out.path <- here::here('output', 'cv')
cv1 <- read_singular_local_out(here::here(out.path, 'default1/'), 'baseline')
cv2 <- read_singular_local_out(here::here(out.path, 'default2/'), 'baseline')
cv3 <- read_singular_local_out(here::here(out.path, 'default3/'), 'baseline')
cv4 <- read_singular_local_out(here::here(out.path, 'default4/'), 'baseline')
cv5 <- read_singular_local_out(here::here(out.path, 'default5/'), 'baseline')

cvlist <- list(cv1, cv2, cv3, cv4, cv5)
cv <- do.call(rbind, cvlist)

# cv simulation batches
cv.simul.files1 <- list.files(here::here('data', 'final_US', 'cross_validation', 'batch1'),
                             pattern = '[0-9]{4}_US_cohort.csv',
                             full.names = TRUE)
raw1 <- do.call(rbind, lapply(cv.simul.files1, read.csv))
cv.simul.files2 <- list.files(here::here('data', 'final_US', 'cross_validation', 'batch2'),
                             pattern = '[0-9]{4}_US_cohort.csv',
                             full.names = TRUE)
raw2 <- do.call(rbind, lapply(cv.simul.files2, read.csv))
cv.simul.files3 <- list.files(here::here('data', 'final_US', 'cross_validation', 'batch3'),
                             pattern = '[0-9]{4}_US_cohort.csv',
                             full.names = TRUE)
raw3 <- do.call(rbind, lapply(cv.simul.files3, read.csv))
cv.simul.files4 <- list.files(here::here('data', 'final_US', 'cross_validation', 'batch4'),
                             pattern = '[0-9]{4}_US_cohort.csv',
                             full.names = TRUE)
raw4 <- do.call(rbind, lapply(cv.simul.files4, read.csv))
cv.simul.files5 <- list.files(here::here('data', 'final_US', 'cross_validation', 'batch5'),
                             pattern = '[0-9]{4}_US_cohort.csv',
                             full.names = TRUE)
raw5 <- do.call(rbind, lapply(cv.simul.files5, read.csv))

rawlist <- list(raw1, raw2, raw3, raw4, raw5)
raw.dat <- do.call(rbind, rawlist)
```

@@

```{r}
# pidp filter
#cv.pidps <- cv %>% select(pidp) %>% unique()
#raw.pidps <- raw %>% select(pidp) %>% unique()
#both.pidps <- intersect(cv.pidps, raw.pidps)

# year filter
cv.years <- cv %>% select(time) %>% unique()
cv.years <- cv.years[['time']]

#cv <- cv.dat %>% filter(pidp %in% both.pidps$pidp)
raw <- raw.dat %>% filter(time %in% cv.years)

#rm(cv.pidps, raw.pidps, both.pidps, cv.years)
rm(cv.years)
```

For all analyses, results will be calculated on each output separately and pooled. For example, the mean hh_income over time will be calculated for each output run seperately, and the mean of these 5 means is what will be plotted. This will be plotted against the entire raw sample.

# hh_income

```{r}
cv.mean.plots(cv1, cv2, cv3, cv4, cv5, raw, 'hh_income')
```

@@

@@

```{r}
snapshot_OP_plots(raw, cv, 'hh_income')
```

@@

# SF_12

```{r}
cv.mean.plots(cv1, cv2, cv3, cv4, cv5, raw, 'SF_12')
```

```{r}
income.pivoted <- combine_and_pivot_long(df1 = cv, 
                                         df1.name = 'simulated', 
                                         df2 = raw, 
                                         df2.name = 'raw', 
                                         var = 'SF_12')

compare_boxplot_long(pivoted.df = income.pivoted,
                     var = 'SF_12',
                     scen1 = 'raw',
                     scen2 = 'simulation')

compare_boxplot_long(pivoted.df = income.pivoted,
                     var = 'SF_12',
                     scen1 = 'raw',
                     scen2 = 'simulation',
                     subset.max = 5000,
                     subset.min = -5000)
```

```{r}
snapshot_OP_plots(raw, cv, 'SF_12')
```

@@

cv_ordinal_plots(pivoted.df = hous.pivoted, 
                 var = 'housing_quality',
                 save = FALSE)
```

cv_ordinal_plots(pivoted.df = nhsafe.pivoted, 
                 var = 'neighbourhood_safety',
                 save = FALSE)
```

cv_ordinal_plots(pivoted.df = lnly.pivoted, 
                 var = 'loneliness',
                 save = FALSE)
```

@@

@@
